A comparison of design-based and model-based approaches for spatial data.

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor grammatical updates

---
title: A comparison of design-based and model-based approaches for spatial data.
author:
  - name: Michael Dumelle
    email: Dumelle.Michael@epa.gov
    affiliation: USEPA
    footnote: 1
  - name: Matt Higham
    email: mhigham@stlaw.edu
    affiliation: STLAW
    footnote: 1
  - name: Lisa Madsen
    affiliation: OSU
  - name: Anthony R. Olsen
    affiliation: USEPA
  - name: Jay M. Ver Hoef
    affiliation: NOAA
address:
  - code: USEPA
    address: United States Environmental Protection Agency, 200 SW 35th St, Corvallis, Oregon, 97333
  - code: STLAW
    address: Saint Lawrence University Department of Math, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
  - code: OSU
    address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
  - code: NOAA
    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  This is the abstract.

@@

We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function. Many common choices for this function assume stationarity and isotropy (CITE Cressie?) with spatial covariance that decreases with increasing Euclidean distance between population units. The primary function used throughout the simulations and applications of this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
\mbox{}
\begin{align}\label{equation:expcov}
\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
\end{align}
where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$, and $\mathbbm{1}$ is an indicator function. However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [pg. 80 - 93][@chiles1999geostatistics].

With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed. While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based. Neither require a particular distribution for $\mathbf{z}$.

<!-- a little weird to talk about inference being based on distributional assumptions and then throw in this sentence about not assuming a distribution for z. -->

@@

For the following simulation results, we simulated 1040 different gridded populations, each of size 900 with sample size 150. For the model-based approach (FPBK), population units were selected via Independent Random Sample. For GRTS, the local mean variance was used.

The response was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1. For model-based, we assumed the correct form of the covariance function (Exponential), but estimated the spatial parameters with REML.

@@

* both good: correctly specified model with high correlation
* break model: highly non-normal errors with small sample size
* break design: small area estimation

@@

<!-- spsurvey is talked about in GRTS. so either move that down here or move this to FPBK. -->

@@

<!-- References from Brus2020statistical, de1990model, Brus1997random -->

>>>>

---
title: A comparison of design-based and model-based approaches for spatial data.
author:
  - name: In alphabetical order Michael Dumelle
    email: Dumelle.Michael@epa.gov
    affiliation: USEPA
    footnote: 1
  - name: Matt Higham
    email: mhigham@stlaw.edu
    affiliation: STLAW
    footnote: 1
  - name: Lisa Madsen
    affiliation: OSU
  - name: Anthony R. Olsen
    affiliation: USEPA
  - name: Jay M. Ver Hoef
    affiliation: NOAA
address:
  - code: USEPA
    address: United States Environmental Protection Agency, 200 SW 35th St, Corvallis, Oregon, 97333
  - code: STLAW
    address: Saint Lawrence University Department of Math, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
  - code: OSU
    address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
  - code: NOAA
    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  This is the abstract.

@@

We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function. It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases. Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\cov(\bm{\delta})$ is
\mbox{}
\begin{align}\label{equation:expcov}
\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
\end{align}
where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$, and $\mathbbm{1}$ is an indicator function. However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp. 80-93].

With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed. While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.

@@

For the following simulation results, we simulated 1040 different gridded populations, each of size 900 (on the unit square) with sample size 150. For the design-based approach, population units were selected via GRTS, the Horvitz-Thompson estimator was used,and the local mean variance was used. For the model-based approach (FPBK), population units were selected via Independent Random Sampling (IRS) and the appropriate prediction and prediction variance formulas were used.

The response was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1. For model-based, we assumed the correct form of the covariance function (exponential), but estimated the spatial parameters with REML.

@@

* both good: correctly specified model with high correlation (we did this in Table \ref{tab:sim_one})
* break model: highly non-normal errors with small sample size
* break design: small area estimation

@@

@@

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor grammatical updates

  \title{A comparison of design-based and model-based approaches for spatial
data.}
    \author[USEPA]{Michael Dumelle\corref{1}}
   \ead{Dumelle.Michael@epa.gov} 
    \author[STLAW]{Matt Higham\corref{1}}
   \ead{mhigham@stlaw.edu} 
    \author[OSU]{Lisa Madsen}

@@

The model-based approach assumes the data are a random realization of a
data-generating process. Randomness is often incorporated through
distributional assumptions on this process. Instead of estimating fixed
but unknown parameters (as in the design-based approach), the goal of
model-based inference in the spatial context is often \emph{prediction}
of an unknown quantity. For example, suppose the realized mean of all
population units is the quantity of interest. Instead of
\emph{estimating} a fixed unknown mean, we are \emph{predicting} the
value of the mean, a random variable. We know that if we sampled all
population units, we would have an exact prediction for the mean of our
one realized process, without any uncertainty. But the true mean of the
spatial process that generated our realized data is still not known.
When predicting the realized mean, we typically are not interested in
the underlying process's true mean.

@@

Figure \ref{fig:fig1} shows

@@

Sampling designs can incorporate spatial locations to obtain samples
that are spatially balanced with respect to the population (Stevens Jr
and Olsen, 2004). A sample is spatially balanced with respect to the
population if the sampled population units are a miniature of the
population units. A sample is a miniature of the population if the
distribution of the sampled population units mirrors the density of all
population units. Spatial balance with respect to the population is
different than spatial balance with respect to geography. A sample that
is spatially balanced with respect to geography is spread out in some
type of equidistant manner over geographical space and is not meant to
be miniatures of the population. When we refer to spatial balance
henceforth, we mean spatial balance with respect to the population.

Spatially balanced samples are useful because they tend to yield
estimates that have lower variance than estimates constructed from
sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
2013; Stevens Jr and Olsen, 2004; Wang et al., 2013). To quantify
spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
based on Voroni polygons. The first spatially balanced sampling
algorithm that saw widespread use was the Generalized Random
Tessellation Stratified (Stevens Jr and Olsen, 2004). Since GRTS was
developed, several other spatially balanced sampling algorithms have
emerged, including the Local Pivotal Method (Grafström et al., 2012;
Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
Halton Iterative Partitioning (Robertson et al., 2018) algorithms. We
focus on the Generalized Random Tessellation Stratified (GRTS) algorithm
to select spatially balanced sampling because the algorithm has several
attractive properties detailed by Stevens Jr and Olsen (2004) and
Dumelle et al. (2021).

The GRTS algorithm is used to sample from finite and infinite
populations and works by utilizing a mapping between two-dimensional and
one-dimensional space. The population units in two-dimensional space are
divided into cells using a hierarchical index. Population units are then
mapped to a one-dimensional line via the hierarchical indexing. The line
length of each population unit equals its inclusion probability. A
systematic sample is conducted on the line and these samples are linked
to a population unit in two-dimensional space, which results in the
desired sample size. Stevens Jr and Olsen (2004) provide and Dumelle et
al. (2021) provide further details. The GRTS algorithm is available in
the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al.,
2021).

After collecting a sample, the data are used to estimate population
parameters. The Horvitz-Thompson estimator (Horvitz and Thompson, 1952)
and its continuous analog (Cordy, 1993) yield unbiased estimates of
population means and totals. For example, if \(\tau\) is a population
total, then the Horvitz-Thompson estimator of \(\tau\) (denoted by
\(\hat{\tau}_{ht}\)), is given by \begin{align}\label{eq:ht}
  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
\end{align} where \(Z_i\) and \(\pi_i\) are the observed value and
inclusion probability of the \(i\)th population unit selected in the
sample. Horvitz and Thompson (1952) and Sen (1953) provide variance
estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
they rely on calculating \(\pi_{ij}\), the probability that population
unit \(i\) and population unit \(j\) are included in the sample, which
can be very difficult to calculate. Second, they ignore the spatial
locations of the population units. To address these drawbacks, Stevens
Jr and Olsen (2003) proposed a local neighborhood variance estimator.
The local neighborhood variance estimator does not rely on \(\pi_{ij}\),
and it incorporates spatial locations by assigning higher weights to
nearby observations. Stevens Jr and Olsen (2003) show this variance
estimators tends to reduce the variability associated with estimating
\(\tau\). This yields confidence intervals for \(\tau\) that are
narrower than confidence intervals constructed from variance estimators
ignoring spatial locations.

@@

Finite Population Block Kriging (FPBK) is an alternative to
sampling-based methods (Ver Hoef, 2008). FPBK expands the geostatistical
kriging framework to the finite population setting. Instead of basing
inference off of a specific sampling design, we assume the data are
generated by a spatial process with parameters that can be estimated
using the framework of a model.

Ver Hoef (2008) gives details on the theory of FPBK, but some of the
basic principles are summarized below. For a response variable
\(\mathbf{z}\) that can be measured on a finite number of \(N\) sites,
we want to predict some linear function of the response variable,
\(\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
\(\mathbf{b}\) is a vector of weights. For example, if we want to
predict the population total across all sites, then we would use a
vector of 1's for the weights.

Typically, however, we only have a sample of the \(N\) sites. Denoting
quantities that are part of the sampled sites with a subscript \emph{s}
and quantities that are part of the unsampled sites with a subscript
\emph{u},

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
design matrices for the sampled and unsampled sites, respectively, and
\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
sampled and unsampled sites. Denoting
\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
that \(E(\bm{\delta})\) = \(\mathbf{0}\).

We also typically assume that there is spatial correlation in
\(\bm{\delta}\), which can be modeled using a covariance function. Many
common choices for this function assume that spatial covariance
decreases with increasing Euclidean distance between sites. The primary
function used throughout the simulations and applications of this
manuscript is the Exponential covariance function: the \(i,j^{th}\)
entry for \(\mathop{\mathrm{{var}}}(\bm{\delta})\) is \mbox{}
\begin{align}\label{equation:expcov}
\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
\end{align} where \(h_{i,j}\) is the distance between sites \(i\) and
\(j\), and \(\bm{\theta}\) is a vector of spatial covariance parameters
of the partial sill \(\theta_1\), the range \(\theta_2\), and the nugget
\(\theta_3\). However, any spatial covariance function could be used in
the place of the Exponential, including functions that allow for
anisotropy {[}pg. 80 - 93{]}(Chiles and Delfiner, 1999).

With the above model formulation, the Best Linear Unbiased Predictor
(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\) and its prediction variance
can be computed. While details of the derivation are in (Ver Hoef,
2008), we note here that the predictor and its variance are both
moment-based. Neither require a particular distribution for
\(\mathbf{z}\).

@@

For the following simulation results, we simulated 1040 different
gridded populations, each of size 900 with sample size 150. For the
model-based approach (FPBK), sites were selected via Independent Random
Sample. For GRTS, the local mean variance was used.

The response was normally distributed with an exponential covariance
function with partial sill of 0.9, effective range of \(\sqrt{2}\), and
a nugget of 0.1. For model-based, we assumed the correct form of the
covariance function (Exponential), but estimated the spatial parameters
with REML.

@@

\begin{itemize}
\tightlist
\item
  both good: correctly specified model with high correlation
\item
  break model: highly non-normal errors with small sample size
\item
  break design: small area estimation
\end{itemize}

@@

FPBK can be readily performed in \texttt{R} with the \texttt{sptotal}
package (Higham et al., 2020). We use \texttt{sptotal} for both the
simulation analysis and the application, estimating parameters with
Restricted Maximum Likelihood (REML).

@@

\leavevmode\hypertarget{ref-cordy1993extension}{}%
Cordy, C.B., 1993. An extension of the horvitz---thompson theorem to
point sampling from a continuous universe. Statistics \& Probability
Letters 18, 353--362.

>>>>

  \title{A comparison of design-based and model-based approaches for spatial
data.}
    \author[USEPA]{In alphabetical order Michael Dumelle\corref{1}}
   \ead{Dumelle.Michael@epa.gov} 
    \author[STLAW]{Matt Higham\corref{1}}
   \ead{mhigham@stlaw.edu} 
    \author[OSU]{Lisa Madsen}

@@

The model-based approach assumes the data are a random realization of a
data-generating process. Randomness is often incorporated through
distributional assumptions on this process. Instead of estimating fixed
but unknown parameters (as in the design-based approach), the goal of
model-based inference in the spatial context is often \emph{prediction}
of an unknown quantity. For example, suppose the realized mean of all
population units is the quantity of interest. Instead of
\emph{estimating} a fixed unknown mean, we are \emph{predicting} the
value of the mean, a random variable. We know that if we sampled all
population units, we would have an exact prediction for the mean of our
one realized process, without any uncertainty. But we are typically not
interested in the true, unknown mean of the underlying process.

@@

Description of Figure \ref{fig:fig1} goes here.

@@

The design-based approach can use spatial locations to obtain spatially
balanced samples. First we discuss spatial balance with respect to the
population (Stevens Jr and Olsen, 2004). A sample is spatially balanced
with respect to the population if the sampled population units are a
miniature of the population units. A sample is a miniature of the
population if the distribution of the sampled population units mirrors
the density of all population units. Spatial balance with respect to the
population is different than spatial balance with respect to geography.
A sample that is spatially balanced with respect to geography is spread
out in some type of equidistant manner over geographical space and is
not meant to be miniatures of the population. When we refer to spatial
balance henceforth, we mean spatial balance with respect to the
population.

Spatially balanced samples are useful because they tend to yield
estimates that have lower variance than estimates constructed from
sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
2013; Stevens Jr and Olsen, 2004; Wang et al., 2013). To quantify
spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
based on Voroni polygons. The first spatially balanced sampling
algorithm that saw widespread use was the Generalized Random
Tessellation Stratified (Stevens Jr and Olsen, 2004). Since GRTS was
developed, several other spatially balanced sampling algorithms have
emerged, including the Local Pivotal Method (Grafström et al., 2012;
Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
Halton Iterative Partitioning (Robertson et al., 2018). We focus on the
Generalized Random Tessellation Stratified (GRTS) algorithm to select
spatially balanced sampling because it has several attractive properties
detailed by Stevens Jr and Olsen (2004) and Dumelle et al. (2021).

The GRTS algorithm is used to sample from finite and infinite
populations and works by utilizing a mapping between two-dimensional and
one-dimensional space. The population units in two-dimensional space are
divided into cells using a hierarchical index. Population units are then
mapped to a one-dimensional line via the hierarchical indexing. The line
length of each population unit equals its inclusion probability. A
systematic sample is conducted on the line and these samples are linked
to a population unit in two-dimensional space, which results in the
desired sample. Stevens Jr and Olsen (2004) provide and Dumelle et al.
(2021) provide further details.

After collecting a sample using the GRTS algorithm, the data are used to
estimate population parameters. The Horvitz-Thompson estimator (Horvitz
and Thompson, 1952) yields unbiased estimates of population means and
totals. For example, if \(\tau\) is a population total, then the
Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
is given by \begin{align}\label{eq:ht}
  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
\end{align} where \(Z_i\) and \(\pi_i\) are the observed value and
inclusion probability of the \(i\)th population unit selected in the
sample. A similar formula exists for estimating the mean, \(\mu\).
Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
for \(\hat{\tau}_{ht}\), but they have two drawbacks. First, they rely
on calculating \(\pi_{ij}\), the probability that population unit \(i\)
and population unit \(j\) are included in the sample, and this can be
very difficult to calculate. Second, they ignore the spatial locations
of the population units. To address these drawbacks, Stevens Jr and
Olsen (2003) proposed a local neighborhood variance estimator. The local
neighborhood variance estimator does not rely on \(\pi_{ij}\), and it
incorporates spatial locations by assigning higher weights to nearby
observations. Stevens Jr and Olsen (2003) show this variance estimators
tends to reduce the estimated standard error of \(\hat{\tau}\), yielding
narrower confidence confidence intervals for \(\tau\).

@@

Finite Population Block Kriging (FPBK) is a model-based approach that
expands the geostatistical Kriging framework to the finite population
setting (Ver Hoef, 2008). Instead of basing inference off of a specific
sampling design, we assume the data are generated by a spatial process.
Ver Hoef (2008) gives details on the theory of FPBK, but some of the
basic principles are summarized below. Let
\(\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}\)
be a response variable that can be measured at the \(N\) population
units and is represented as an \(N \times 1\) vector. Suppose we want to
predict some linear function of the response variable,
\(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where \(\mathbf{b}\)
is a \(1 \times N\) vector of weights. For example, if we want to
predict the population total across all population units, then we would
use a vector of 1's for the weights.

Typically, however, we only have a sample of the \(N\) population units.
Denoting quantities that are part of the sampled population units with a
subscript \emph{s} and quantities that are part of the unsampled
population units with a subscript \emph{u},

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
design matrices for the sampled and unsampled population units,
respectively; \(\beta\) is the parameter vector of fixed effects; and
\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
sampled and unsampled population units, respectively. Denoting
\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
the expectation of \(\bm{\delta}\) equals \(\mathbf{0}\).

We also typically assume that there is spatial correlation in
\(\bm{\delta}\), which can be modeled using a covariance function. It is
common to assume the covariance function is second-order stationary and
isotropic (Cressie, 1993), and that the spatial covariance decreases as
the separation between population units increases. Many spatial
covariance functions exist, but the primary function we use throughout
the simulations and applications in this manuscript is the exponential
covariance function: the \(i,j^{th}\) entry for
\(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
\begin{align}\label{equation:expcov}
\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
\end{align} where \(h_{i,j}\) is the distance between population units
\(i\) and \(j\), and \(\bm{\theta}\) is a vector of spatial covariance
parameters of the partial sill \(\theta_1\), the range \(\theta_2\), and
the nugget \(\theta_3\), and \(\mathbbm{1}\) is an indicator function.
However, any spatial covariance function could be used in the place of
the exponential, including functions that allow for non-stationarity or
anisotropy (Chiles and Delfiner, 1999, pp. 80--93).

With the above model formulation, the Best Linear Unbiased Predictor
(BLUP) for \(f(\mathbf{b}'\mathbf{z})\) and its prediction variance can
be computed. While details of the derivation are in (Ver Hoef, 2008), we
note here that the predictor and its variance are both moment-based.

@@

For the following simulation results, we simulated 1040 different
gridded populations, each of size 900 (on the unit square) with sample
size 150. For the design-based approach, population units were selected
via GRTS, the Horvitz-Thompson estimator was used,and the local mean
variance was used. For the model-based approach (FPBK), population units
were selected via Independent Random Sampling (IRS) and the appropriate
prediction and prediction variance formulas were used.

The response was normally distributed with an exponential covariance
function with partial sill of 0.9, effective range of \(\sqrt{2}\), and
a nugget of 0.1. For model-based, we assumed the correct form of the
covariance function (exponential), but estimated the spatial parameters
with REML.

@@

\begin{itemize}
\tightlist
\item
  both good: correctly specified model with high correlation (we did
  this in Table \ref{tab:sim_one})
\item
  break model: highly non-normal errors with small sample size
\item
  break design: small area estimation
\end{itemize}

@@

The GRTS algorithm and the local neighborhood variance estimator are
available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
et al., 2021). FPBK can be readily performed in \texttt{R} with the
\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
for both the simulation analysis and the application, estimating
parameters with Restricted Maximum Likelihood (REML).

@@
